\section{Competitive Equilibrium \& Welfare}

\subsection{Market Equilibrium}
$Q^D(p) = Q^S(p)$. With inverse demand $P^D(Q)$ and inverse supply $P^S(Q)$:
$$P^D(Q^*) = P^S(Q^*)$$

\subsection{Welfare}
$CS = \int_0^{Q^*} P^D(Q)\,dQ - P^*Q^*$. \quad $PS = P^*Q^* - \int_0^{Q^*} P^S(Q)\,dQ$.

$W = CS + PS$. Competitive equilibrium maximizes $W$ (First Welfare Theorem).

\subsection{Tax Incidence (Partial Equilibrium)}
Per-unit tax $t$. Let $\epsilon_D < 0$, $\epsilon_S > 0$:
$$\frac{dp^D}{dt} = \frac{\epsilon_S}{\epsilon_S - \epsilon_D}, \qquad \frac{dp^S}{dt} = \frac{\epsilon_D}{\epsilon_D - \epsilon_S}$$
The more \textbf{inelastic} side bears more. Incidence is independent of statutory assignment.

\textit{Application:} Payroll tax ``paid by employers'' vs.\ ``paid by workers'' has the same incidence: what matters is $\epsilon_S$ vs.\ $\epsilon_D$, not statutory labels.

\textbf{Deadweight loss} (Harberger triangle):
$$DWL \approx \frac{1}{2}\frac{\epsilon_S |\epsilon_D|}{\epsilon_S - \epsilon_D} \frac{t^2}{P}Q$$
DWL grows with $t^2$ $\implies$ broad low taxes dominate narrow high taxes.

\textbf{Key implications:}
\begin{enumerate}[nosep]
  \item Doubling tax rate \textit{quadruples} DWL
  \item Two 5\% taxes create $\approx \frac{1}{2}$ the DWL of one 10\% tax
  \item Tax most heavily goods with inelastic demand/supply
  \item DWL exists even with inelastic labor supply if non-cash compensation (fringe benefits) is untaxed---the tax distorts the \textit{form} of compensation
\end{enumerate}

\textbf{Pigouvian correction:} Tax $t = MC_{\text{external}}$ restores efficiency. Unlike ordinary taxes, this \textit{reduces} DWL. Target the externality directly---tax emissions, not electricity (see \S Externalities).

\subsection{Non-Price Market Clearing}
When prices cannot adjust:
\begin{itemize}[nosep]
    \item \textbf{Wait time:} Kidney market --- excess demand cleared by queue. Wait time $\tau$ acts as a ``shadow price'': $P_{eff} = P_{ceiling} + v\tau$
    \item \textbf{Search costs:} Min wage --- effective leisure price $p_\ell = w - p_s$. Ratio $p_\ell/p_f < A \implies$ less food, more leisure, plus DWL from search
    \item \textbf{Quality adjustment:} Rent control --- landlords reduce maintenance until effective price matches equilibrium
\end{itemize}

\textit{Modeling:} Replace constrained price with shadow price $P_{eff} = P_{\text{cap}} + \text{non-price cost}$, then re-solve.


